

\begin{thebibliography}{99}

\bibitem[Illich, 1973]{Illich} Illich I. (1973). \textit{Tools for conviviality}. (dans le domaine public, disponible sur Internet)

\bibitem[Foucault, 1975]{Foucault} Foucault M. (1975)  \textit{Surveiller et punir : Naissance de la prison}.  Gallimard

\bibitem[Deleuze, 1986]{Deleuze} Deleuze G. (1986). \textit{Foucault}. Les Éditions de Minuit.

\bibitem[Qian, Lehman, 2017]{Qian_Lehman_2017} 
Qian Y., Lehman J. (2017). Students’ Misconceptions and Other Difficulties in Introductory Programming: A Literature Review. \textit{ACM Trans. Comput. Educ. 18, 1, Article 1 (October 2017)} \url{https://dl.acm.org/doi/10.1145/3077618}

\bibitem[Wenger, 1998]{wenger} 
Wenger, E. (1998).\textit{Communities of practice: Learning, meaning, and identity}. Cambridge University Press. \url{https://doi.org/10.1017/CBO9780511803932}

\bibitem{Papert, 1980} Papert, S. (1980). \textit{Mindstorms: Children, Computers, and Powerful Ideas}. Basic Books.

\bibitem[Messer, 2023]{messer2023} 
Messer, M., Brown, N.C.C., Kölling, M., \& Shi, M. (2023). Automated Grading 
and Feedback Tools for Programming Education: A Systematic Review. \textit{arXiv preprint} 
arXiv:2306.11722.

\bibitem[Zimmermann, 2024]{zimmermann2024} Zimmermann, A. E., King, E. E., \& Bose, D. D. (2024). Effectiveness 
and Utility of Flowcharts on Learning in a Classroom Setting: A Mixed-Methods Study. 
\textit{American Journal of Pharmaceutical Education, 88}(1), 100591.


\bibitem[Crews, Zielger, 98]{CrewsZielger98}Crews, T., Ziegler, U. (1998) The flowchart interpreter for introductory programming courses, FIE '98. \textit{28th Annual Frontiers in Education Conference. Moving from 'Teacher-Centered' to 'Learner-Centered' Education. Conference Proceedings (Cat. No.98CH36214), Tempe, AZ, USA, 1998, pp. 307-312 vol.1}, \url{doi: 10.1109/FIE.1998.736854}.


\bibitem[Sentance,2017]{Sentance2017WiPSCE}
Sentance, S., \& Waite, J. (2017).
PRIMM: Exploring pedagogical approaches for teaching text-based programming in school.
In \textit{Proceedings of the 12th Workshop on Primary and Secondary Computing Education (WiPSCE '17)} (pp.~113--114).
New York, NY: ACM.
https://doi.org/10.1145/3137065.3137084

\bibitem[Sentance, 2019]{Sentance2019CSE}
Sentance, S., Waite, J., \& Kallia, M. (2019).
Teaching computer programming with PRIMM: A sociocultural perspective.
\textit{Computer Science Education}, \textit{29}(2--3), 136--176.
\url{https://doi.org/10.1080/08993408.2019.1608781}

\bibitem[Anderson, Krathwohl,2001]{AndersonKrathwohl2001}
Anderson, L. W., \& Krathwohl, D. R. (Eds.). (2001).
\textit{A taxonomy for learning, teaching, and assessing: A revision of Bloom's taxonomy of educational objectives}.
New York, NY: Longman.

\bibitem[Krathwohl, 2002]{Krathwohl2002} Krathwohl, D. R. (2002). A revision of Bloom's taxonomy: An overview. \textit{Theory Into Practice}, \textit{41}(4), 212--218. \url{https://doi.org/10.1207/s15430421tip4104_2}

\bibitem[Fuller, 2007]{Fuller2007}
Fuller, U., Johnson, C. G., Ahoniemi, T., Cukierman, D., Hernan-Losada, I., Jackova, J., Lahtinen, E., Lewis, T. L., McGee Thompson, D., Riedesel, C., \& Thompson, E. (2007).
Developing a computer science-specific learning taxonomy.
\textit{ACM SIGCSE Bulletin}, \textit{39}(4), 152--170.
\url{https://doi.org/10.1145/1345375.1345438}

\bibitem[Hattie, Timperley, 2007]{HattieTimperley2007}
Hattie, J., \& Timperley, H. (2007).
The power of feedback.
\textit{Review of Educational Research}, \textit{77}(1), 81--112. \url{https://doi.org/10.3102/003465430298487}

\bibitem[Shute, 2008]{Shute2008}
Shute, V. J. (2008).
Focus on formative feedback.
\textit{Review of Educational Research}, \textit{78}(1), 153--189. \url{https://doi.org/10.3102/0034654307313795}

\bibitem[Black, Wiliam, 1998]{BlackWiliam1998}
Black, P., \& Wiliam, D. (1998).
Inside the black box: Raising standards through classroom assessment.
\textit{Phi Delta Kappan}, \textit{80}(2), 139--148.

\bibitem[Parker, Garcia, Kao, 2022]{ACES} Parker, M., Garcia, L., Kao, Y. (2022). A Pair of ACES: An Analysis of Isomorphic Questions on an Elementary Computing Assessment.
\textit{Association for Computing Machinery} 
\url{https://doi.org/10.1145/3501385.3543979}

\bibitem[Sovietov, 2022]{sovietov} 
Sovietov, P. (2022). Automatic Generation of Programming Exercises. \textit{Institute of Information Technologies MIREA – Russian technological university Moscow, Russia - sovetov@mirea.ru} 10.48550/arXiv.2205.11304. 

\bibitem[Keuning, 2018]{Keuning2018}
Keuning, H., Jeuring, J., \& Heeren, B. (2018).
A systematic literature review of automated feedback generation for programming exercises.
\textit{ACM Transactions on Computing Education}, \textit{19}(1), 1--43.
https://doi.org/10.1145/3231711

\bibitem[Altman, Musk, 2015]{email} Révélations mises au jour lors du procès opposant Elon Musk à OpenAi et son PDG (Sam Altman):
\begin{quote}
    \textbf{On Jun 24, 2015, at 10:24 AM, Sam Altman wrote:}

1) The mission would be to create the first general Al and use it for individual empowerment—ie, the distributed version of the future that seems the safest. More generally, safety should be a first-class requirement.
\end{quote} Cité dans la presse en novembre 2024. \url{https://www.businessinsider.com/emails-between-sam-altman-elon-musk-kicked-off-openai-2024-11} (consulté en août 2025).

\bibitem[www.didask.com]{didask} Didask. (n.d.). \textit{Qu'est-ce qu'un outil auteur ? - Le guide complet}. Recopié de \url{https://www.didask.com (consulté en 2025)}.

\bibitem[Doc officielle Pyodide]{pyodide} Pyodide Development Team. (2023). \textit{Pyodide Documentation}. Récupéré de \url{https://pyodide.org} (consulté en 2024).

\bibitem[Doc officielle CodeMirror]{codemirror6} Haverbeke, M. (2021). \textit{CodeMirror 6 Documentation}. Récupéré de \url{https://codemirror.net/6/} (consulté en 2024).

\bibitem[Doc officielle CodeMirror]{ace} Ace Editor. (n.d.). \textit{Ace - The High Performance Code Editor for the Web}. Récupéré de \url{https://ace.c9.io} (consulté en 2024, finalement abandonné au profit de CodeMirror).

\bibitem[Doc officielle Skulp]{skulpt} \url{https://skulpt.org/} (consulté en 2024, finalement abandonné au profit de Pyodide uniquement).

\bibitem[Doc officielle Brython]{brython} \url{https://brython.info/} (consulté en 2024, finalement abandonné au profit de Pyodide uniquement).

\bibitem[Doc offielle Python ast]{python-ast} Python Software Foundation. (2023). \textit{Python Documentation: ast --- Abstract Syntax Trees}. Extrait de \url{https://docs.python.org/3/library/ast.html} (consulté en 2024).

\bibitem[Doc offielle Python dis]{python-dis} Python Software Foundation. (2023). \textit{Python Documentation: dis --- Disassembler for Python bytecode}. Extrait de \url{https://docs.python.org/3/library/dis.html} (consulté en 2024).

\bibitem[Kluyver, 2012]{greentreesnakes} Kluyver, T. (2012). \textit{Green Tree Snakes: Getting to and from ASTs}. Documentation en ligne, \url{https://greentreesnakes.readthedocs.io} (consulté en 2024).

\bibitem[Zeller, 2022]{fuzzingbook} Zeller, A. et al. (2022). \textit{The Fuzzing Book} – Chapitre "Fuzzing with Grammars". En ligne, \url{https://www.fuzzingbook.org} (consulté en 2024).

\bibitem[Page GitHub de Rahul Gopinath]{rahulgopinath} Gopinath, R. \textit{Page personnelle}. En ligne, \url{https://github.com/rahulgopinath/rahulgopinath.github.io} (consulté en 2024)

\bibitem[Flowchart For\_Each]{for_each_exemple} Discussion sur Stackoverflow. \url{https://stackoverflow.com/questions/20580028/flowchart-for-each-loop-loop-without-variable-increment}

\end{thebibliography}

